# Week 3

## Dictionary-based techniques

An extension of word frequency analyses, which we covered last week, are so-called "dictionary-based" techniques. In their most basic form, these analyses use an index of target terms and classify the corpus of interest based on their presence or absence. The technical dimensions of this type of analysis are covered in the chapter section by @krippendorff_content_2004, and some of the issues attending them in the article by - @loughran_when_2011.

We will also be reading two examples of the application of these techniques by @martins_rise_2020 and @young_affective_2012. Here, we will be discussing how successful the authors are in measuring the phenomenon of interest ("prosociality" and "tone" respectively). Questions about sampling and representativeness will again be relevant here, and will naturally inform our assessments of this work.

Questions:

1. Are *general* dictionaries possible; or do they have to be domain-specific?
2. How do we know if our dictionary is accurate?
3. How could we enhance/supplement dictionary-based techniques?

**Required reading**:

- @martins_rise_2020
- @young_affective_2012

- @loughran_when_2011
- @krippendorff_content_2004 (pp.283-289)

**Further reading**:

- @tausczik_psychological_2010
- @brier_computer_2011
- @barbera_automated_2021

**Slides**:

- Week 3 [Slides](https://docs.google.com/presentation/d/1rgYCYGtZ7resCd7oVsGnaCpwWKduunl8s1famjtGtBY/edit?usp=sharing)

### Happy words

In this section, we'll have a quick overview of how we're processing text data when conducting basic sentiment analyses.

We'll first load the packages we need. 

```{r, message=F, warning=F}
library(stringi)
library(dplyr)
library(tidytext)
library(ggplot2)
library(scales)
```

As we discussed in the lectures, we might find in our text of the class's collective thoughts that there was an increase in "happy" words over time. 

I have simulated a dataset of text split by weeks, students, and words plus whether or not the word is the word "happy" where `0` means it is not the word "happy" and `1` means it is. 

```{r, echo = F}
## Varying total happy word numbers example

lipsum_text <- data.frame(text = stri_rand_lipsum(5000, start_lipsum = TRUE))

lipsum_text$happyn <- as.character("")
lipsum_text$happyu <- as.character("")
lipsum_text$happyd <- as.character("")

lipsum_text$week <- as.integer(rep(seq.int(1:10), 5000/10))
lipsum_text$student <- sample(1:30,5000, replace = T)

for(i in 1:nrow(lipsum_text)) {
  week <- lipsum_text[i,5]
  happyn <- paste(rep("happy", times = sample(1:100, 1)), collapse = " ")
  happyu <- paste(rep("happy", times = sample(1:100, 1)*week), collapse = " ")
  happyd <- paste(rep("happy", times = sample(1:100, 1)*1/week), collapse = " ")
  lipsum_text[i, 2] <- happyn
  lipsum_text[i, 3] <- happyu
  lipsum_text[i, 4] <- happyd
}

lipsum_text$happyn <- paste(lipsum_text$text, lipsum_text$happyn)
lipsum_text$happyu <- paste(lipsum_text$text, lipsum_text$happyu)
lipsum_text$happyd <- paste(lipsum_text$text, lipsum_text$happyd)

happyn <- lipsum_text %>% 
  select(happyn, week, student) %>%
  unnest_tokens(word, happyn) %>%
  group_by(week, student) 
happyu <- lipsum_text %>% 
  select(happyu, week, student) %>%
  unnest_tokens(word, happyu) %>%
  group_by(week, student) 
happyd <- lipsum_text %>% 
  select(happyd, week, student) %>%
  unnest_tokens(word, happyd) %>%
  group_by(week, student)

happyn$happy <- as.integer(grepl("happy", x = happyn$word))
happyu$happy <- as.integer(grepl("happy", x = happyu$word))
happyd$happy <- as.integer(grepl("happy", x = happyd$word))

happys <- list(happyn, happyu, happyd)

```

We have three datasets: one with a constant number of "happy" words; one with an increasing number of "happy" words; and one with a decreasing number of "happy" words. These are called: `happyn`, `happyu`, and `happyd` respectively.

```{r}
head(happyn)
```
```{r}
head(happyu)
```
```{r}
head(happyd)
```

We can then see the trend in "happy" words over by week and student.

First, the dataset where we have a constant number of happy words over time.

```{r}
happyn %>%
  group_by(week, student) %>%
  mutate(index_total = n()) %>%
  filter(happy == 1) %>%
  summarise(
    sum_hap = sum(happy),
    index_total = index_total,
    prop_hap = sum_hap / index_total
  ) %>%
  distinct() %>%
  ggplot() +
  geom_point(aes(week, prop_hap)) +
  geom_smooth(aes(week, prop_hap)) +
  scale_x_continuous(breaks = pretty_breaks())
```

And now the simulated data with an increasing number of happy words.

```{r}
happyu %>%
  group_by(week, student) %>%
  mutate(index_total = n()) %>%
  filter(happy == 1) %>%
  summarise(
    sum_hap = sum(happy),
    index_total = index_total,
    prop_hap = sum_hap / index_total
  ) %>%
  distinct() %>%
  ggplot() +
  geom_point(aes(week, prop_hap)) +
  geom_smooth(aes(week, prop_hap)) +
  scale_x_continuous(breaks = pretty_breaks())
```

And finally a decreasing number of happy words.

```{r}
happyd %>%
  group_by(week, student) %>%
  mutate(index_total = n()) %>%
  filter(happy == 1) %>%
  summarise(
    sum_hap = sum(happy),
    index_total = index_total,
    prop_hap = sum_hap / index_total
  ) %>%
  distinct() %>%
  ggplot() +
  geom_point(aes(week, prop_hap)) +
  geom_smooth(aes(week, prop_hap)) +
  scale_x_continuous(breaks = pretty_breaks())
```

### Normalizing sentiment

But as we discussed in the lecture, we also know that just because the total number of happy words increases, this isn't indication on its own that we're getting happier as a class over time. 

Before we can begin to make any such inference, we need to normalize by the total number of words each week. 

Below, I simulate data where the number of happy words is actually the same each week (our `happyn` dataset above). I join these data with three other datasets: `happylipsumn`, `happylipsumu`, and `happylipsumd`. These are datasets of random text, with the *same* number of happy words. The first of these also has the *same* number of total words each week. The second two, however, have *differing* number of *total* words each week: `happylipsumu` has an increasing number of total words each week; `happylipsumd` has a decreasing number of total words each week.

```{r, echo=F}
lipsum_text$lipsumn <- as.character("")
lipsum_text$lipsumu <- as.character("")
lipsum_text$lipsumd <- as.character("")

for(i in 1:nrow(lipsum_text)) {
  week <- lipsum_text[i,5]
  lipsumn <- paste(rep(stri_rand_lipsum(1, start_lipsum = TRUE), 
                       times = 1), collapse = " ")
  lipsumu <- paste(rep(stri_rand_lipsum(1, start_lipsum = TRUE), 
                       times = sample(1:10,1)*week), collapse = " ")
  lipsumd <- paste(rep(stri_rand_lipsum(1, start_lipsum = TRUE), 
                       times = sample(1:10,1)*1/week), collapse = " ")
  lipsum_text[i, 7] <- lipsumn
  lipsum_text[i, 8] <- lipsumu
  lipsum_text[i, 9] <- lipsumd
}

lipsum_text$happylipsumn <- paste(lipsum_text$lipsumn, lipsum_text$happyn)
lipsum_text$happylipsumu <- paste(lipsum_text$lipsumu, lipsum_text$happyn)
lipsum_text$happylipsumd <- paste(lipsum_text$lipsumd, lipsum_text$happyn)

happylipsumn <- lipsum_text %>% 
  select(happylipsumn, week, student) %>%
  unnest_tokens(word, happylipsumn) %>%
  group_by(week, student) 
happylipsumu <- lipsum_text %>% 
  select(happylipsumu, week, student) %>%
  unnest_tokens(word, happylipsumu) %>%
  group_by(week, student) 
happylipsumd <- lipsum_text %>% 
  select(happylipsumd, week, student) %>%
  unnest_tokens(word, happylipsumd) %>%
  group_by(week, student) 

happylipsumn$happy <- as.integer(grepl("happy", x = happylipsumn$word))
happylipsumu$happy <- as.integer(grepl("happy", x = happylipsumu$word))
happylipsumd$happy <- as.integer(grepl("happy", x = happylipsumd$word))

lipsums <- list(happylipsumn, happylipsumu, happylipsumd)

```

Again, as you see below, we're splitting by week, student, word, and whether or not it is a "happy" word. 

```{r}
head(happylipsumn)

head(happylipsumu)

head(happylipsumd)
```

Then if we plot the number of happy words *divided* by the number of total words each week for each student in each of these datasets, we get the below. 

```{r, echo = F, message = F, warning = F}

library(gridExtra)

plots=list()

for (i in seq_along(lipsums)) {
  
  lipsum <- lipsums[[i]]
  
  plots[[i]] <- lipsum %>%
    group_by(week, student) %>%
    mutate(index_total = n()) %>%
    filter(happy==1) %>%
    summarise(sum_hap = sum(happy),
              index_total = index_total,
              prop_hap = sum_hap/index_total) %>%
    distinct() %>%
    ggplot() +
    geom_point(aes(week, prop_hap)) +
    geom_smooth(aes(week, prop_hap)) +
    scale_x_continuous(breaks= pretty_breaks()) +
      theme(axis.text.y=element_text(size=20),
            axis.text.x=element_text(size=20),
            axis.title.x=element_text(size=20, face="bold"),
            axis.title.y=element_text(size=20, face="bold"))
}

do.call("grid.arrange", c(plots, ncol = 2))

```